---
title       : Class01
subtitle    : example code used in DM class01
author      : 
job         : 
framework   : shower        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---
  
## Install R packages
  
R has a lot of powerful packages. You need to load a package into memory before using it.

```r
library(ggplot2) ## a package for plotting
```

--- .modal
  
## Install R packages

If you see an error message:
```
there is no package called 'ggplot2'
```
it means you don't have this package and you need to install it.
* Using GUI
* Using command

--- .modal

## Install R packages (using command)

```{r eval=FALSE, tidy=TRUE}
## install directly from the default repository http://cran.r-project.org/
install.packages('ggplot2') 

## download the package source, and then install from the downloaded file 
download.file('http://cran.r-project.org/src/contrib/ggplot2_0.9.3.1.tar.gz',
             destfile='~/Downloads/ggplot2_0.9.3.1.tar.gz')
install.packages('~/Downloads/ggplot2_0.9.3.1.tar.gz',repo=NULL,type='source')
```

---  .modal

## Getting help

```{r eval=FALSE}
?read.delim # Access a function's help file

??base::delim # Search for 'delim' in all help files for functions in 'base'

help.search("delim") # Search for 'delimited' in all help files

RSiteSearch("parsing text") # Search for the term 'parsing text' on the R site
```

---  .modal

## Working with data

```r
## give the filename of your data; the file can be an URL
data.file = 'http://www.yurulin.com/class/spring2014_datamining/data/ml_hackers/01_heights_weights_genders.csv'
heights.weights = read.csv(data.file , header = TRUE , sep = ',')
View(heights.weights)
```

```{r}
class(heights.weights)
```

---  .modal

## Working with data

```{r}
dim(heights.weights) # show the dimension of the data
```

---  .modal

## Working with data: summary

```{r, results='markup'}
## R summary function
# ?summary
summary(heights.weights)
```

---  .modal

## Working with data: summary

```{r}
summary(heights.weights$Height) # summary for the 'Height' column  
```

---  .modal

## Working with data

```{r}
## Means, Medians, Modes
heights = heights.weights$Height
mean(heights)

median(heights)

min(heights); max(heights)
```


---  .modal

## Working with data: Quantiles

```{r}
quantile(heights)

quantile(heights, prob = 0.25)

quantile(heights, probs = seq(0, 1, by = 0.20))
```

---  .modal

## Working with data

```{r}
## Standard Deviations and Variances
sd(heights)

var(heights)

## Tabulation
table(heights.weights$Gender)
```

---  .modal

## Working with data

```{r}
## Correlation 
weights = heights.weights$Weight
cor(heights,weights)
```

---  .modal

## Plotting

```{r, fig.width=4, fig.height=4}
## Histogram and Density
library('ggplot2') 
theme_set(theme_bw()) # change the default ggplot theme to black-and-white

ggplot(heights.weights, aes(x = Height)) +
  geom_histogram(binwidth=1) 
```

---  .modal

## Plotting

```{r, fig.width=4, fig.height=4}
ggplot(heights.weights, aes(x = Height)) +
  geom_histogram(binwidth=0.5) # try different value for binwidth 
```

---  .modal

## Plotting

```{r, fig.width=4, fig.height=4}
ggplot(heights.weights, aes(x = Height)) +
  geom_density()
```

---  .modal

## Plotting

```{r, fig.width=5, fig.height=4}
## conditional density plot
ggplot(heights.weights, aes(x = Weight, fill = Gender)) +
  geom_density(alpha = 0.5) 
```

---  .modal

## Plotting

```{r, fig.width=4, fig.height=4}
ggplot(heights.weights, aes(x = Weight)) +
  geom_density(alpha = 0.5) + facet_grid(Gender ~ .) 
```

---  .modal

## Plotting

```{r, fig.width=5, fig.height=4}
ggplot(heights.weights, aes(x = Weight, fill = Gender)) +
  geom_density(alpha = 0.5) + facet_grid(Gender ~ .)  
```

---  .modal

## Plotting

```{r, fig.width=4, fig.height=4}
## Scatterplot
ggplot(heights.weights, aes(x = Weight, y = Height)) + geom_point()
```

---  .modal

## Plotting

```{r, fig.width=5, fig.height=4}
ggplot(heights.weights, aes(x = Weight, y = Height, color = Gender)) + geom_point()
```

---  .modal

## Plotting

```{r, fig.width=4, fig.height=4}
ggplot(heights.weights, aes(x = Weight, y = Height)) + geom_point(shape=1) + # use hollow circles 
  geom_smooth(method=lm, # add linear regression line
              se=FALSE) # (by default includes 95% confidence region)  
```

--- 

## More about plotting

See: [http://www.cookbook-r.com/Graphs/](http://www.cookbook-r.com/Graphs/)

--- .modal

## Example 1

```{r}
require(lattice)  
require(nutshell) 
data(births2006.smpl)
births2006.smpl[1:3,]
```

--- .modal

## Example 1

```{r}
dim(births2006.smpl) # the dimension of the data frame (data matrix)
```


--- .modal

## Example 1

```{r}
## generate frequency tables
births.dow=table(births2006.smpl$DOB_WK)
births.dow 
```

--- .modal

## Example 1

```{r, fig.height=4, fig.width=4}
barchart(births.dow,ylab="Day of Week",col="black")
```


--- .modal

## Example 1

```{r, fig.height=4, fig.width=4, warning=FALSE}
## plot by ggplot
ggplot(data = births2006.smpl, aes(x=factor(DOB_WK))) + 
  geom_histogram(stat="count", binwidth=1) + 
  scale_x_discrete("Day of Week") + scale_y_continuous("Freq") + 
  coord_flip()
```

--- .modal

## Example 1

```{r}
dob.dm.tbl=table(WK=births2006.smpl$DOB_WK,MM=births2006.smpl$DMETH_REC)
dob.dm.tbl
```


--- .modal

## Example 1

```{r}
dob.dm.tbl=dob.dm.tbl[,-2]
dob.dm.tbl
```

--- .modal

## Example 1

```{r, fig.height=4, fig.width=5, message=TRUE, warning=FALSE}
## plot by ggplot
ggplot(data = births2006.smpl, aes(x=factor(DOB_WK),fill=factor(DMETH_REC) )) + 
  geom_histogram(stat="count") + 
  scale_x_discrete("Day of Week") + scale_y_continuous("Freq") +
  scale_fill_discrete(name="method of\ndelivery") +
  coord_flip()
```


--- .modal

## Example 1

```{r, fig.height=4, fig.width=4}
## generate conditional histogram and density plots
histogram(~DBWT|DPLURAL,data=births2006.smpl,layout=c(1,5),col="black")
```


--- .modal

## Example 1

```{r, fig.height=4, fig.width=4, message=TRUE, warning=FALSE}
## plot by ggplot
ggplot(data = births2006.smpl,aes(x=DBWT)) + facet_grid(DPLURAL~.,scales="free_y") + 
  geom_histogram(binwidth=500)
```


--- .modal

## Example 1

```{r, fig.height=5, fig.width=5}
histogram(~DBWT|DMETH_REC,data=births2006.smpl,layout=c(1,3),col="black")
```


--- .modal

## Example 1

```{r, fig.height=5, fig.width=5}
densityplot(~DBWT|DPLURAL,data=births2006.smpl,layout=c(1,5),plot.points=FALSE,col="black")
```


--- .modal

## Example 1

```{r, fig.height=5, fig.width=5, warning=FALSE}
## plot by ggplot
ggplot(data = births2006.smpl,aes(x=DBWT)) + facet_grid(DPLURAL~., scales="free_y") + 
  geom_density()  
```

--- .modal

## Example 2: thin-tailed v. heavy-tailed

```{r}
set.seed(1)
normal.values <- rnorm(100000, 0, 1)
cauchy.values <- rcauchy(100000, 0, 1)
range(normal.values)

range(cauchy.values)  
```


--- .modal

## Example 2: thin-tailed v. heavy-tailed

```{r, fig.height=5, fig.width=6, warning=FALSE}
D=data.frame(X = cauchy.values,distribution='cauchy')
D=rbind(D,data.frame(X = normal.values,distribution='normal'))
ggplot(data=D, aes(x=X, linetype=distribution)) + 
  geom_density() + 
  scale_x_continuous(limits = c(-6, 6))
```

--- .modal

## Example 3: Power law

```{r}
## the frequency of family names can be downloaded from the 2000 Census:
## http://www.census.gov/genealogy/www/data/2000surnames/
data.url = 'http://www.yurulin.com/class/archive/fall2013_datamining/data/others'
data.file = sprintf('%s/Top1000names.csv',data.url)
topfamilynames = read.csv(data.file,stringsAsFactors=FALSE)
namecount = topfamilynames[-1,1:3]
names(namecount) = c('Surnames','Rank','Count')
D=namecount[order(as.numeric(namecount$Count),decreasing=T),]
D[1:3,]
```


--- .modal

## Example 3: Power law

```{r, fig.height=4, fig.width=4}
data = as.numeric(namecount$Count)
# rank the family names according to frequency, from largest to smallest
data = sort(data,decreasing=TRUE) 
hist(data,xlab='Number of Family Surnames',main='Histogram of Family Surnames')
```


--- .modal

## Example 3: Power law

```{r, fig.height=4, fig.width=4}
plot(data,type='l',
     ylab='Frequency of Family Surnames',
     xlab='Index',
     main='Plot of Family Surnames')
```


--- .modal

## Example 3: Power law

```{r, fig.height=4, fig.width=4}
plot(data,type='p',log='xy',
     ylab='Frequency of Family Surnames',
     xlab='Index',
     main='Log-log Plot of Family Surnames')  
```

