---
title       : Class10
subtitle    : Network Mining
author      : Yu-Ru Lin
job         : 
framework   : shower        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
toc         : true
toc_depth   : 2

--- #toc
## Class10
  
* [Set up](#set-up)
* [network representation](#rep)
* [network centrality](#cen)
* [network layout](#layout)
* [network community](#comm)
* [graph with attributes](#attr)

--- #set-up .modal 

## Install R packages
```{r, eval=FALSE}
## this tutorial uses the following packages
```

--- #rep .sscode-nowrap 
## network representation

* Network mining using igraph
* adjacency matrices and adjacency lists
* directed and undirected networks

```{r}
library(igraph)

## create a toy matrix
m = matrix(nrow=3,ncol=3)
m[1,1]=0
m[1,2]=1
m[1,3]=1
m[2,1]=1
m[2,2]=0
m[2,3]=0
m[3,1]=0
m[3,2]=1
m[3,3]=0
m
```

--- .scode-nowrap .compact
## network representation
```{r, dependson=c(-1)}
lab = c('a','b','c') ## label nodes as 'a','b','c'
## create directed graph from the toy matrix
g = graph.adjacency(m,mode="directed") 
g
```

--- .scode-nowrap .compact
## network representation
```{r, dependson=c(-1)}
set.seed(1) ## reset random seed before drawing
plot(g,vertex.label=lab,edge.color='black')
```

---  .scode-nowrap .compact
## network representation
```{r, dependson=c(-1)}
adj = get.adjacency(g)
adj
```

---  .scode-nowrap .compact
## network representation
```{r, dependson=c(-1)}
el = get.edgelist(g)
el
```

---  .scode-nowrap 
## network representation
```{r, dependson=c(-1)}
degree(g)
degree(g, mode="in")
degree(g, mode="out")
colSums(m) ## in-degree
rowSums(m) ## out-degree
```

---  .scode-nowrap .compact
## network representation
```{r, dependson=c(-1)}
## how far apart?
shortest.paths(g, mode='out')
diameter(g)
average.path.length(g,directed=T) ## 8/6
```

---  .scode-nowrap .compact
## network representation
```{r, dependson=c(-1)}
## how dense are they?
graph.density(g) ## 4/6

## connected components of a graph
clusters(g)
```

---  .scode-nowrap .compact
## network representation
```{r, dependson=c(-1)}
## create undirected graph from the same matrix
g = graph.adjacency(m,mode="undirected") 
g
```

---  .scode-nowrap .compact
## network representation
```{r, dependson=c(-1)}
set.seed(1)
plot(g,vertex.label=lab,edge.color='black')
```

---  .scode-nowrap .compact
## network representation
```{r, dependson=c(-1)}
adj = get.adjacency(g)
adj
```

---  .scode-nowrap .compact
## network representation
```{r, dependson=c(-1)}
el = get.edgelist(g)
el
```

--- #cen .sscode-nowrap
## network centrality

* degree centrality
* closeness centrality
* betweenness centrality
* PageRank
* HITS (authority and hub scores)

```{r, dependson=c(-1)}
## KARATE network.
## Zachary's karate club: social network of friendships between 34 members of a karate club at a US university in the 1970s. Please cite W. W. Zachary, An information flow model for conflict and fission in small groups, Journal of Anthropological Research 33, 452-473 (1977).

## load the famous karate network from igraph package
g = graph.famous("Zachary")
g
```

--- .scode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
set.seed(1)
E(g)$color = "black"
plot(g)
```

--- .scode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
## obtain node degrees
deg = degree(g)
deg
top = order(deg,decreasing=T)[1:5] ## the top-5 nodes with highest degrees
top 
```

--- .sscode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
## size node by degree
V(g)$size = abs(deg) * 1.5
V(g)$label.color = "black"
E(g)$color = "black"
V(g)[ top ]$label.color = "red" ## highlight the top-5 nodes
set.seed(1)
plot(g)
title("degree centrality")
```

--- .scode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
## compute node closeness centrality
clo = closeness(g)
clo
```

--- .sscode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
top = order(clo,decreasing=T)[1:5]

## size node by closeness
V(g)$size = abs(clo) * 1000
V(g)$label.color = "black"
V(g)[ top ]$label.color = "red" ## highlight the top-5 nodes
set.seed(1)
plot(g)
title("closeness")
```

--- .scode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
## compute node betweenness centrality
bet = betweenness(g)
bet
```

--- .sscode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
top = order(bet,decreasing=T)[1:5]

## size node by betweenness
V(g)$size = abs(bet) * 0.1
V(g)$label.color = "black"
V(g)[ top ]$label.color = "red" ## highlight the top-5 nodes
set.seed(1)
plot(g)
title("betweenness")
```

--- .scode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
pg = page.rank(g)$vector
pg
```

--- .sscode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
top = order(pg,decreasing=T)[1:5]

## size node by pagerank
V(g)$size = abs(pg) * 250
V(g)$label.color = "black"
V(g)[ top ]$label.color = "red" ## highlight the top-5 nodes
set.seed(1)
plot(g)
title("PageRank")
```

--- .sscode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
## compute HITS
aut = authority.score(g)$vector
top = order(aut,decreasing=T)[1:5]

## size node by authority scores
V(g)$size = abs(aut) * 20
V(g)$label.color = "black"
V(g)[ top ]$label.color = "red" ## highlight the top-5 nodes
set.seed(1)
plot(g)
title("HITS (authority)")
```

--- .sscode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
hub = hub.score(g)$vector 
top = order(hub,decreasing=T)[1:5]

## size node by hub scores
V(g)$size = abs(aut) * 20
V(g)$label.color = "black"
V(g)[ top ]$label.color = "red" ## highlight the top-5 nodes
set.seed(1)
plot(g)
title("HITS (hub)")
```

--- .scode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
## apply HITS on the toy matrix (directed network)
m
lab = c('a','b','c') ## label nodes as 'a','b','c'
## create directed graph from the toy matrix
g = graph.adjacency(m,mode="directed") 
g
```

--- .sscode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
aut = authority.score(g)$vector

## size node by authority scores
V(g)$size = abs(aut) * 30
V(g)$label.color = "black"
set.seed(1)
plot(g,edge.color='black')
title("HITS (authority)")
```

--- .sscode-nowrap .compact
## network centrality
```{r, dependson=c(-1)}
hub = hub.score(g)$vector 

## size node by hub scores
V(g)$size = abs(hub) * 30
V(g)$label.color = "black"
set.seed(1)
plot(g,edge.color='black')
title("HITS (hub)")
```

--- #layout .scode-nowrap .compact
## network layout
```{r, dependson=c(-1)}
## load the famous karate network from igraph package
g = graph.famous("Zachary")
E(g)$color = "black"

set.seed(1)
plot(g, layout=layout.random, main='random')
```

--- .scode-nowrap .compact
## network layout
```{r, dependson=c(-1)}
set.seed(1)
plot(g, layout=layout.kamada.kawai, main='kamada kawai (KK)')
```

--- .scode-nowrap .compact
## network layout
```{r, dependson=c(-1)}
set.seed(1)
plot(g, layout=layout.circle, main = 'circle')
```

--- .scode-nowrap .compact
## network layout
```{r, dependson=c(-1)}
set.seed(1)
plot(g, layout=layout.fruchterman.reingold, main = 'fruchterman reingold (FR)')
```

--- .scode-nowrap .compact
## network layout
```{r, dependson=c(-1)}
set.seed(1)
plot(g, layout=layout.lgl, main = 'lgl')
```

--- .scode-nowrap .compact
## network layout
```{r, dependson=c(-1)}
set.seed(1)
plot(g, layout=layout.graphopt, main = 'graphopt')
```

--- #comm .scode-nowrap 
## network community
 * modularity
 * random walk
 * spin-glass
 * eigenvector
 * ...
```{r, dependson=c(-1)}
g = graph.famous("Zachary")

## get random-walk community
wc = walktrap.community(g)
modularity(wc)
membership(wc)
```

--- .scode-nowrap .compact
## network community
```{r, dependson=c(-1)}
set.seed(1)
plot(wc, g, main='random-walk community')
```

--- .scode-nowrap .compact
## network community
```{r, dependson=c(-1)}
dend = as.dendrogram(wc, use.modularity=TRUE)
plot(dend, nodePar=list(pch=c(NA, 20)))
```

--- .scode-nowrap .compact
## network community
```{r, dependson=c(-1)}
dendPlot(wc)
```

--- .scode-nowrap .compact
## network community
```{r, dependson=c(-1)}
## get modularity-based community
fc = fastgreedy.community(g)
modularity(fc)
membership(fc)
```

--- .scode-nowrap .compact
## network community
```{r, dependson=c(-1)}
set.seed(1)
plot(fc, g, main='modularity community')
```

--- .scode-nowrap .compact
## network community
```{r, dependson=c(-1)}
dendPlot(fc)
```

--- .scode-nowrap .compact
## network community
```{r, dependson=c(-1)}
sg = spinglass.community(g)
modularity(sg)
set.seed(1)
plot(sg, g, main='spin-glass community')
```

--- .scode-nowrap .compact
## network community
```{r, dependson=c(-1)}
le = leading.eigenvector.community(g)
modularity(sg)
set.seed(1)
plot(le, g, main='eigenvector community')
```

--- .scode-nowrap .compact
## network community
```{r, dependson=c(-1)}
## compare the simlarity between two community assignment
## @see http://hosho.ees.hokudai.ac.jp/~kubo/Rdoc/library/igraph/html/compare.communities.html
compare(sg, le, method="rand")
compare(sg, le, method="nmi")
compare(membership(sg), membership(le), method="rand")
```

--- .scode-nowrap .compact
## network community
```{r, dependson=c(-1)}
## highlight between-community edges
le = leading.eigenvector.community(g,steps=1)
le
```

--- .ssscode-nowrap .compact
## network community
```{r, dependson=c(-1)}
V(g)$color = ifelse(le$membership==1, "lightblue", "green")

scale <- function(v, a, b) {
  v = v-min(v) ; v = v/max(v) ; v = v * (b-a) ; v+a
}

V(g)$size = scale(sapply(le$eigenvector,abs), 10, 20)
E(g)$color = "grey"
E(g)[ V(g)[ color=="lightblue" ] %--% V(g)[ color=="green" ] ]$color = "red"
set.seed(1)
plot(g, layout=layout.kamada.kawai, vertex.label.font=2)
```

--- .scode-nowrap .compact #attr
## graph with attributes
```{r, dependson=c(-1)}
## @see https://github.com/igraph/igraph/tree/master/doc/presentations/wien08
traits = read.csv("http://cneurocvs.rmki.kfki.hu/igraphbook/traits.csv", head=FALSE)
traits[1:3,]
```

--- .scode-nowrap .compact
## graph with attributes
```{r, dependson=c(-1)}
rel = read.csv("http://cneurocvs.rmki.kfki.hu/igraphbook/relations.csv", head=FALSE)
rel[1:3,]
```

--- .sscode-nowrap .compact
## graph with attributes
```{r, dependson=c(-1)}
g = graph.empty(directed=T)
g = add.vertices(g, nrow(traits), 
                 name=as.character(traits[,1]), age=traits[,2],
                 gender=as.character(traits[,3]))
V(g)$name
V(g)$age
V(g)$gender
```

--- .scode-nowrap .compact
## graph with attributes
```{r, dependson=c(-1)}
names = sapply(strsplit(V(g)$name, " "), "[",1)
V(g)$name = names
ids = (1:length(names)) # index starts from 1
names(ids) = names
ids
```

--- .sscode-nowrap .compact
## graph with attributes
```{r, dependson=c(-1)}
from = as.character(rel[,1])
to = as.character(rel[,2])
edges = matrix(c(ids[from], ids[to]), nc=2)
edges
```

--- .scode-nowrap .compact
## graph with attributes
```{r, dependson=c(-1)}
g = add.edges(g, t(edges), 
              room=as.character(rel[,3]),
              friend=rel[,4], advice=rel[,5])
g
```

--- .scode-nowrap .compact
## graph with attributes
```{r, dependson=c(-1)}
male = which(V(g)$gender=='M')
V(g)$shape = "circle"
V(g)$shape[male] ="rectangle"
E(g)$color = "black"
E(g)[ room=="Y" ]$color = "red"
set.seed(1)
plot(g, layout=layout.kamada.kawai, edge.color=E(g)$color)
```

--- .scode-nowrap .compact
## graph with attributes
```{r, dependson=c(-1)}
E(g)$color = "black"
E(g)[ advice>2 ]$color = "green"
set.seed(1)
plot(g, layout=layout.kamada.kawai, edge.color=E(g)$color)
```

--- .scode-nowrap .compact
## graph with attributes
```{r, dependson=c(-1)}
E(g)$color = "black"
E(g)[ friend>4 ]$color = "skyblue"
set.seed(1)
plot(g, layout=layout.kamada.kawai, edge.color=E(g)$color)
```

