---
title       : Class12
subtitle    : Data Processing and Features
author      : Yu-Ru Lin
job         : 
framework   : shower        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
toc         : true
toc_depth   : 2

--- #toc
## Class12
  
* [Set up](#set-up)
* [Big data set](#big)
* [Sparse matrix](#sparse)
* [Feature selection](#featureselection)
* [Frequent patterns](#freq)

--- #set-up .modal 

## Install R packages
```{r, eval=FALSE}
## this tutorial uses the following packages
install.packages('sqldf')
```

--- .scode-nowrap .compact #big
## Big data set

* When dealing with big data files, do not read everything into memory.
* We use the sample data from: http://grouplens.org/datasets/hetrec-2011/
* See: http://files.grouplens.org/datasets/hetrec2011/hetrec2011-movielens-readme.txt
  
```{r}
## give the input filename
data.path = 'hetrec2011'
ifilename = sprintf('%s/movies.dat',data.path)
## read the first n=3 lines
lines = readLines(ifilename,n=3)
lines
```

--- .ssscode-nowrap .compact 
## Big data set
```{r, dependson=c(-1)}
## read the first 3 rows into a data.frame object
df = read.csv(ifilename, sep='\t', nrows=3) ## we saw from the first few lines that the field separator is <TAB>
df
```

--- .ssscode-nowrap .compact 
## Big data set
```{r, dependson=c(-1)}
## read row 101--103
df = read.csv(ifilename, sep='\t', nrows=3, skip=101, header=F) ## if we set 'header=T', the first non-skip row will be treated as header (which is incorrect)
df
```

--- .scode-nowrap .compact 
## Big data set
```{r, dependson=c(-1)}
## count how many lines in the file
# length(readLines(ifilename)) ## not recommended because it reads the file in memory
system(sprintf("wc -l %s",ifilename), intern=T ) 
## you can use 'wc' command on unix-based machine (linux, mac, etc.)

suppressWarnings(suppressPackageStartupMessages( ## avoid printing warnings
  require(R.utils)
)) ## assumig you have installed R.utils
countLines(ifilename)
```

--- .scode .compact 
## Big data set
```{r, dependson=c(-1)}
suppressWarnings(suppressPackageStartupMessages( ## avoid printing warnings
  require(sqldf)
)) ## assumig you have installed sqldf
read.csv.sql(ifilename, "select count(*) from file", sep='\t')
```

--- .scode-nowrap .compact 
## Big data set
```{r, dependson=c(-1)}
sql = 'select id,title,year from file limit 1' ## read 1 row
df = read.csv.sql(ifilename, sql, sep='\t')
df
```

--- .scode-nowrap .compact 
## Big data set
```{r, dependson=c(-1)}
sql = 'select id,title,year from file limit 10' ## read 10 rows
df = read.csv.sql(ifilename, sql, sep='\t')
df
```

--- .scode-nowrap .compact 
## Big data set
```{r, dependson=c(-1)}
sql = 'select id,title,year from file limit 5,10' 
## skip the first 5 rows, and read the next 10 rows
df = read.csv.sql(ifilename, sql, sep='\t')
df
```

--- .scode-nowrap .compact 
## Big data set
```{r, dependson=c(-1)}
sql = 'select year,count(*) from file group by year limit 10' 
## you can do more if you are familiar with SQL 
df = read.csv.sql(ifilename, sql, sep='\t')
df
```

* if you have mysql database, you can use 'RMySQL' package.
* Other tips (using 'fread'): http://davetang.org/muse/2013/09/03/handling-big-data-in-r/


--- .scode-nowrap .compact #sparse
## Sparse matrix
When dealing with big matrix, use sparse format to construct matrix.

```{r, dependson=c(-1)}
ifilename = sprintf('%s/movies.dat',data.path)
movie.df = read.csv(ifilename, sep='\t')
dim(movie.df)
```

--- .scode-nowrap .compact 
## Sparse matrix
```{r, dependson=c(-1)}
movie.df[1:3,1:4] ## check a small part
```

--- .scode-nowrap .compact 
## Sparse matrix
```{r, dependson=c(-1)}
ifilename = sprintf('%s/%s',data.path, 'movie_genres.dat')
m2g.df = read.csv(ifilename, sep='\t', header=T)
m2g.df[1:3,]
dim(m2g.df)
```

--- .scode-nowrap .compact 
## Sparse matrix
```{r, dependson=c(-1)}
m2g.df[1:10,] ## check a small part
```

--- .scode-nowrap .compact 
## Sparse matrix
```{r, dependson=c(-1)}
## create a numeric index for genres, so that we can create a matrix
g = unique(as.character(m2g.df$genre))
length(g) ## how many genres?
m = unique(as.character(m2g.df$movieID))
length(m) ## how many movies?
idx2g = match(as.character(m2g.df$genre),g)
idx2g[1:20]; length(idx2g)
idx2m = match(as.character(m2g.df$movieID),m)
idx2m[1:20]; length(idx2m)
```

--- .scode-nowrap .compact 
## Sparse matrix
```{r, dependson=c(-1)}
require(Matrix);require(MASS)
## construct a movie-to-genre sparse matrix; the format is (i,j,v), meaning entry(i,j)=v
m2g = sparseMatrix(i=idx2m,j=idx2g,x=rep(1,nrow(m2g.df)) )
dim(m2g)
object.size(m2g)
print(object.size(m2g),units="Mb")  ## gives you size of the object in Mb

## report current memory used by R, e.g., __ Mb in my machine
# print(object.size(x=lapply(ls(), get)), units="Mb") 
denseM2G = as.matrix(m2g) ## dense matrix
print(object.size(denseM2G),units="Mb") 

## report current memory used by R, e.g., __ Mb in my machine
print(object.size(x=lapply(ls(), get)), units="Mb") 

rm(denseM2G)
## report current memory used by R, e.g., __ Mb in my machine
print(object.size(x=lapply(ls(), get)), units="Mb")
```

--- .scode-nowrap .compact 
## Sparse matrix
```{r, dependson=c(-1)}
m2g[1:3,]
```

--- .scode-nowrap .compact 
## Sparse matrix
```{r, dependson=c(-1)}
```

--- .scode-nowrap .compact 
## Sparse matrix
```{r, dependson=c(-1)}
```

