---
title       : Class03
subtitle    : binary classification
author      : 
job         : 
framework   : shower        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
--- #toc

## Class03
* [Set up](#set-up)
* [Split data into two classes](#two-class)
* [Logistic Regression](#logistic-regression)
* [kNN](#knn)

--- #set-up .modal 

## Install R packages
```{r eval=FALSE}
## this tutorial uses the following packages
```

--- #two-class .scode-nowrap .compact 

## Split data into two classes
```{r, fig.height=4.5, fig.width=8}
library(ggplot2) ## use ggplot for plotting
data.url = 'http://www.yurulin.com/class/spring2014_datamining/data/ml_hackers/'
data.file = file.path(data.url , '01_heights_weights_genders.csv')
heights.weights <- read.csv(data.file , header = TRUE , sep = ',')
# visualized their gender as the color of each point
ggplot(heights.weights, aes(x = Weight , y = Height , color = Gender)) +
  geom_point()
```

--- .scode-nowrap .compact 

## Split data into two classes

```{r, fig.height=4, fig.width=4, dependson=c(-1)}
heights.weights = transform(heights.weights , Male = ifelse(Gender == 'Male', 1, 0))
heights.weights [1:3,]
```

--- #logistic-regression .model
## Logistic Regression
### Example 1: Death penelty data
   * objective: to identify whether the death penalty more likely if the victim is white, or the crime is more aggravating
   * response variable: getting death panelty or not
   * explanatory variables: severity of the crime, race of victim (black: 0; white: 1), etc

--- .scode-nowrap .compact 
## Logistic Regression
```{r}
## analyzing individual observations
data.url = 'http://www.yurulin.com/class/spring2014_datamining/data/data_text'
dpen = read.csv(sprintf("%s/DeathPenalty.csv",data.url))
dpen[1:4,]
```


--- .scode-nowrap .compact 
## Logistic Regression
```{r, dependson=c(-1), fig.height=4, fig.width=4}
m1=glm(Death~VRace+Agg,family=binomial(link='logit'),data=dpen)
m1
```


--- .scode-nowrap .compact 
## Logistic Regression
```{r, dependson=c(-1), fig.height=4, fig.width=4}
summary(m1)
```


--- .scode-nowrap .compact 
## Logistic Regression
```{r, dependson=c(-1), fig.height=4, fig.width=4}
## calculating logits
exp(m1$coef[2])
exp(m1$coef[3])
```

--- .sscode-nowrap .compact
## Logistic Regression

```{r, dependson=c(-1), fig.height=4, fig.width=4}
## plotting probability of getting death penalty as a function of aggravation
## separately for black (in black) and white (in red) victim
n.dots = 501 ## number of dots in the curves (the nubmer is set to make the curves smooth)
x.ag=dim(n.dots) ## x-value is the aggravating level (severity of the crime)
y.black=dim(n.dots) ## y-value for the black victim curve
y.white=dim(n.dots) ## y-value for the white victim curve
for (i in 1:n.dots) {
  x.ag[i]=(99+i)/100
  is.white = 0
  y.black[i]=exp(m1$coef[1]+m1$coef[2]*is.white+m1$coef[3]*x.ag[i])/(1+exp(m1$coef[1]+m1$coef[2]*is.white+m1$coef[3]*x.ag[i]))
  is.white = 1
  y.white[i]=exp(m1$coef[1]+m1$coef[2]*is.white+m1$coef[3]*x.ag[i])/(1+exp(m1$coef[1]+m1$coef[2]*is.white+m1$coef[3]*x.ag[i]))
}
plot(y.black~x.ag,type="l",col="black",ylab="Prob[Death]",xlab="Aggravation",ylim=c(0,1),main="red line for white victim; black line for black victim")
points(y.white~x.ag,type="l",col="red")
```

--- .scode-nowrap .compact 
## Logistic Regression
```{r, dependson=c(-1), fig.height=4, fig.width=4}
## another format for logistic regression:
## summarized data
dpenother <- read.csv(sprintf("%s/DeathPenaltyOther.csv",data.url))
dpenother
```


--- .scode-nowrap .compact 
## Logistic Regression
```{r, dependson=c(-1), fig.height=4, fig.width=4}
m1=glm(Death~VRace+Agg,family=binomial,weights=Freq,data=dpenother)
m1
```
--- .scode-nowrap .compact 
## Logistic Regression
```{r, dependson=c(-1), fig.height=4, fig.width=4}
summary(m1)
exp(m1$coef[2])
exp(m1$coef[3])
```

