---
title       : Class11
subtitle    : Recommendation
author      : Yu-Ru Lin
job         : 
framework   : shower        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
toc         : true
toc_depth   : 2

--- #toc
## Class11
  
* [Set up](#set-up)
* [Data manipulation](#data)
* [Sample dataset](#sample)
* [Recommender System](#recommender)

--- #set-up .modal 

## Install R packages
```{r, eval=FALSE}
## this tutorial uses the following packages
install.packages('recommenderlab')
```

--- .compact
## Recommendation

* The sample code is based on the examples given in http://cran.r-project.org/web/packages/recommenderlab/vignettes/recommenderlab.pdf
* See information about the installation and more details about what the package can do in the manual.

```{r}
library(recommenderlab)
```

--- .scode-nowrap .compact #data
## Data manipulation

```{r, dependson=c(-1)}
## create a toy reating matrix
m = matrix(sample(c(as.numeric(0:5), NA), 50,
                  replace=TRUE, prob=c(rep(.4/6,6),.6)), ncol=10,
           dimnames=list(user=paste("u", 1:5, sep=''),
                         item=paste("i", 1:10, sep='')))
m
```

--- .scode-nowrap .compact 
## Data manipulation

```{r, dependson=c(-1)}
r = as(m, "realRatingMatrix") ## store the matrix in sparse format
## The realRatingMatrix can be coerced back into a matrix which is identical to the original matrix
identical(as(r, "matrix"),m)
```

--- .sscode-nowrap .compact
## Data manipulation

```{r, dependson=c(-1)}
## a list of users
as(r, "list")
```

--- .sscode-nowrap .compact
## Data manipulation

```{r, dependson=c(-1)}
## rating data in data.frame format
head(as(r, "data.frame"))
```

--- .scode-nowrap .compact
## Data manipulation

```{r, dependson=c(-1)}
## normalization: remove rating bias by subtracting the row mean from all ratings in the row
r_m = normalize(r)
r_m
```

--- .scode-nowrap .compact
## Data manipulation

```{r, dependson=c(-1)}
r_z = normalize(r,method="Z-score")
r_z
```

--- .scode-nowrap .compact
## Data manipulation

```{r, dependson=c(-1)}
## visualize the matrices
image(r, main = "Raw Ratings")
```

--- .scode-nowrap .compact
## Data manipulation

```{r, dependson=c(-1)}
image(r_m, main = "Normalized Ratings")
```

--- .scode-nowrap .compact
## Data manipulation

```{r, dependson=c(-1)}
image(r_z, main = "Normalized Ratings (Z-score)")
```

--- .sscode-nowrap .compact
## Data manipulation

```{r, dependson=c(-1)}
## Binarization: transform the real-valued matrix into a 0-1 matrix based on a user specified threshold (min_ratings)
r_b = binarize(r, minRating=4)
as(r_b, "matrix")
```

--- .scode-nowrap .compact #sample
## Sample dataset

```{r, dependson=c(-1)}
## load the Jester5k data: it contains a sample of 5000 users from the anonymous ratings data from the Jester Online Joke Recommender System collected between April 1999 and May 2003 (Goldberg, Roeder, Gupta, and Perkins, 2001).

data(Jester5k)
Jester5k
```

--- .sscode-nowrap .compact
## Sample dataset

```{r, dependson=c(-1)}
set.seed(1) ## reset random seed before sampling

## use only a subset of the data containing a sample of 1000 users
r = sample(Jester5k, 1000)

## the ratings for the first user
rowCounts(r[1,])
as(r[1,], "list")
```

--- .scode-nowrap .compact
## Sample dataset

```{r, dependson=c(-1)}
rowMeans(r[1,]) ## The user has rated 100 jokes, the list shows the ratings and the user's rating average is 1.4731
```

--- .scode-nowrap .compact
## Sample dataset

```{r, dependson=c(-1)}
## generate a histogram to show the distribution of the ratings
hist(getRatings(r), breaks=100)
```

--- .scode-nowrap .compact
## Sample dataset

```{r, dependson=c(-1)}
## the distribution after normalization
hist(getRatings(normalize(r)), breaks=100)
```

--- .scode-nowrap .compact
## Sample dataset

```{r, dependson=c(-1)}
hist(getRatings(normalize(r, method="Z-score")), breaks=100)
```

--- .scode-nowrap .compact
## Sample dataset

```{r, dependson=c(-1)}
## how many jokes each user has rated
hist(rowCounts(r), breaks=50)
```

--- .scode-nowrap .compact
## Sample dataset

```{r, dependson=c(-1)}
## the mean rating for each joke
hist(colMeans(r), breaks=20)
```

--- .scode-nowrap .compact #recommender
## Recommender System

```{r, dependson=c(-1)}
## (training) create a recommender which generates recommendations solely on the popularity of items (the number of users who have the item in their profile)
r = Recommender(Jester5k[1:1000], method = "POPULAR")
r
## get the model by getModel()
names(getModel(r))
```

--- .sscode-nowrap .compact
## Recommender System

```{r, dependson=c(-1)}
getModel(r)$topN

## (prediction) create top-5 recommendation lists for two users who were not used to learn the model
recom = predict(r, Jester5k[1001:1002], n=5)
recom

as(recom, "list")
```

--- .scode-nowrap .compact
## Recommender System

```{r, dependson=c(-1)}
```

--- .scode-nowrap .compact
## Recommender System

```{r, dependson=c(-1)}
```

