---
title       : Class06
subtitle    : clustering
author      : Yu-Ru Lin
job         : 
framework   : shower        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
toc         : true
toc_depth   : 2

--- #toc
## Class06
  
* [Set up](#set-up)
* [k-means](#kmeans)
* [Hierarchical clustering](#hclu)
* [Clustering evaluation](#eval)

--- #set-up .modal 

## Install R packages
```{r, eval=FALSE}
## this tutorial uses the following packages
```

--- .scode-nowrap .compact #kmeans
## k-means
Protein consumption data

```{r}
## load the data:
data.url = 'http://www.yurulin.com/class/spring2014_datamining/data/data_text'
food = read.csv(sprintf("%s/protein.csv",data.url))
dim(food)

food[1:3,] ## take a look at the first three rows
```

--- .scode-nowrap .compact
## k-means
```{r, dependson=c(-1)}
## k-means clustering on just Red and White meat, with k=3 clusters
set.seed(1) ## fix the random seed to produce the same results 
grpMeat = kmeans(food[,c("WhiteMeat","RedMeat")], centers=3, nstart=10)
grpMeat
```

--- .sscode-nowrap .compact
## k-means
```{r, dependson=c(-1)}
## list the cluster assignments
o=order(grpMeat$cluster)
data.frame(food$Country[o],grpMeat$cluster[o])
```

--- .scode-nowrap .compact
## k-means
```{r, dependson=c(-1), fig.width=5, fig.height=5}
## generate a scatter plot with cluster assignments
plot(food$Red, food$White, type="n", xlim=c(3,19), xlab="Red Meat", ylab="White Meat")
text(x=food$Red, y=food$White, labels=food$Country, col=grpMeat$cluster+1)
```

--- .sscode-nowrap .compact
## k-means
```{r, dependson=c(-1)}
## same analysis, but now with clustering on all nine features;
## change the number of clusters to 7
set.seed(1)
grpProtein = kmeans(food[,-1], centers=7, nstart=10) 
o=order(grpProtein$cluster)
data.frame(food$Country[o],grpProtein$cluster[o])
```

--- .scode-nowrap .compact
## k-means
```{r, dependson=c(-1), fig.width=5, fig.height=5}
## generate a scatter plot for the new result
plot(food$Red, food$White, type="n", xlim=c(3,19), xlab="Red Meat", ylab="White Meat")
text(x=food$Red, y=food$White, labels=food$Country, col=rainbow(7)[grpProtein$cluster])
```

--- .scode-nowrap .compact #hclu
## Hierarchical clustering 
Protein consumption data

```{r}
## use library 'cluster'
library(cluster)
## load the data
data.url = 'http://www.yurulin.com/class/spring2014_datamining/data/data_text'
food = read.csv(sprintf("%s/protein.csv",data.url))
food[1:3,] ## take a look at the first three rows
```

--- .scode-nowrap .compact 
## Hierarchical clustering 
```{r, dependson=c(-1)}
## consider just Red and White meat clusters;
## use the program 'agnes' in the package cluster;
## argument:
## diss=FALSE indicates that we use the dissimilarity matrix that is being calculated from raw data;
## metric="euclidian" indicates that we use Euclidian distance;
## no standardization is used as the default (stand = FALSE);
## use average linkage by default (method = "average"); 
food2 = food[,c("WhiteMeat","RedMeat")]
food2agg = agnes(food2, diss=FALSE, metric="euclidian")
food2agg
```

--- .scode-nowrap .compact 
## Hierarchical clustering 
```{r, dependson=c(-1), fig.width=7, fig.height=5}
plot(food2agg, which.plots=2) ## dendrogram
```

--- .sscode-nowrap .compact 
## Hierarchical clustering 
```{r, dependson=c(-1)}
food2agg$merge  ## describes the sequential merge steps
```

--- .scode-nowrap .compact 
## Hierarchical clustering 
```{r, dependson=c(-1), fig.width=7, fig.height=5}
## identical result obtained by first computing the distance matrix
food2aggv = agnes(daisy(food2), metric="euclidian", diss=T)
plot(food2aggv, which.plots=2)
```

--- .scode-nowrap .compact 
## Hierarchical clustering 
```{r, dependson=c(-1), fig.width=7, fig.height=5}
## use all nine features with Euclidean distance and average linkage
foodagg = agnes(food[,-1],diss=FALSE,metric="euclidian")
plot(foodagg, which.plots=2)  ## dendrogram
```

--- .scode-nowrap .compact 
## Hierarchical clustering 
```{r, dependson=c(-1), fig.width=7, fig.height=5}
# foodagg$merge  ## describes the sequential merge steps

## Using data on all nine variables (features)
## Euclidean distance and single linkage
foodaggsin=agnes(food[,-1],diss=FALSE,metric="euclidian",method="single")
plot(foodaggsin, which.plots=2)	## dendrogram
```

--- .ssscode-nowrap .compact 
## Hierarchical clustering 
```{r, dependson=c(-1), fig.width=7, fig.height=5}
# foodaggsin$merge	## describes the sequential merge steps

## Euclidean distance and complete linkage
foodaggcomp=agnes(food[,-1],diss=FALSE,metric="euclidian",method="complete")
plot(foodaggcomp, which.plots=2)	## dendrogram
```

--- .scode-nowrap .compact 
## Hierarchical clustering 
```{r, dependson=c(-1)}
# foodaggcomp$merge	## describes the sequential merge steps

## use hclust,cutree for hierarchical clustering
data.dist = dist(food[,-1]) ## use dist to obtain distance matrix
hc = hclust(data.dist,method='complete')
data.dist
```

--- .scode-nowrap .compact 
## Hierarchical clustering 
```{r, dependson=c(-1), fig.width=7, fig.height=5}
plot(hc)
```

--- .scode-nowrap .compact 
## Hierarchical clustering 
```{r, dependson=c(-1)}
hc1 = cutree(hc,k=3)
print(hc1)
```

--- .scode-nowrap .compact #eval
## Clustering evaluation
* purity and entropy
* an "elbow" in the plot of SSE against cluster solutions
```{r}
cluster.purity <- function(clusters, classes) {
  sum(apply(table(classes, clusters), 2, max)) / length(clusters)
}

cluster.entropy <- function(clusters,classes) {
  en <- function(x) {
    s = sum(x)
    sum(sapply(x/s, function(p) {if (p) -p*log2(p) else 0} ) )
  }
  M = table(classes, clusters)
  m = apply(M, 2, en)
  c = colSums(M) / sum(M)
  sum(m*c)
}
```

--- .scode-nowrap .compact 
## Clustering evaluation
```{r, dependson=c(-1)}
## test the purity and entropy functions
n = 20
classes = sample(3, n, replace=T) ## randomly assign classes
classes
```

--- .scode-nowrap .compact 
## Clustering evaluation
```{r, dependson=c(-1)}
clusters = sample(2, n, replace=T) ## randomly assign clusters
clusters
```

--- .scode-nowrap .compact 
## Clustering evaluation
```{r, dependson=c(-1)}
table(classes, clusters) ## confusion matrix
```

--- .scode-nowrap .compact 
## Clustering evaluation
```{r, dependson=c(-1)}
cluster.purity(clusters,classes)  ## purity
```

--- .scode-nowrap .compact 
## Clustering evaluation
```{r, dependson=c(-1)}
cluster.entropy(clusters,classes) ## entropy
```

--- .scode-nowrap .compact 
## Clustering evaluation
```{r, dependson=c(-1)}
data.url = 'http://www.yurulin.com/class/spring2014_datamining/data/data_text'
food = read.csv(sprintf("%s/protein.csv",data.url))


wssplot <- function(data, nc=15, seed=1234) {
  wss = dim(nc)
  for (i in 2:nc) {
    set.seed(seed)
    wss[i] = sum(kmeans(data, centers=i)$withinss)
  }
  plot(1:nc, wss, type="b", xlab="Number of Clusters", ylab="Within groups sum of squares")
}
```

--- .scode-nowrap .compact 
## Clustering evaluation
```{r, dependson=c(-1)}
n.clusters = 10
wssplot(scale(food[,-1]),nc=n.clusters)
```
